Autonomous Oceanographic Data Collection Estimating Excursion Sets of Vector-valued Gaussian Random Fields

### 2020-06-23 main_new.tex
updates

%The methods we develop are fairly general, and not limited to the special case of two-dimensional random fields. Hence, the rest of this section will consider generic excursion sets of gaussian random fields with an arbitrary number of output dimensions.
% Say we have an underlying phenomenon that is modeled as a $\no$-variate gaussian random field $\gp$ on some domain $\domain$, 
For the general setting of a $\no$-variate random field, 
we are interested in recovering the set of locations $\es$ in the domain for which the components of $\gp$ lie in some set of specified values $\T\subset \mathbb{R}^{\no}$, in other words \textit{the pre-image of $T$ by $\gp$}:
$$
\es:=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
$$
%
%Our goal here is to develop approaches to quantify and improve the
%characterization of uncertainties on $\es$. \textcolor{red}{Alread mention how observations are included.}

If we assume that $\gp$ has continuous trajectories 
%(almost surely \kc{This seems very handwavy; would delete}) 
and $T$ is closed, then
$\es$ becomes a Random Closed Set \citep{Molchanov2005} and concepts from the theory of random sets will prove useful to study $\es$. 
%under some measure defined on the domain.
%\textcolor{red}{Is this really needed?} 
Note that while some aspects of the developed approaches do not call for a
specific form of $\T$, we will often stick for simplicity to the case
of orthants
($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where
$t_1,\dots, t_{\no} \in \R$) as this will allow efficient calculation
of several key quantities. Note that changing some $\leq$ inequalities
to $\geq$ ones would lead to immediate adaptations.

@@

%Given the observations from $n$ batches, one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$. The mean update formulae is
%It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
% delivers that
% 
%\begin{equation}\label{eq:meanCoK}
%\mu_{[n+1]}(\bm{x})=\mu_{[n]}(\bm{x})+\lambda_{[n+1,n+1]}(\bm{x})^T (\mathbf{z}_{n+1}-\mu(\bm{x}_{n+1})),
%\end{equation}
%where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional sub-vector extracted from
%$\lambda_{[n+1]}(\bm{x})$ that corresponds to the Kriging weigths associated with the last $q_{n+1}$ responses when 
%predicting at $\bm{x}$ relying on all measurements until batch $(n+1)$.
%\text{th}$ batch.
%, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
%The updated co-Kriging residual cross-covariance function is
%\begin{eqnarray}\label{eq:varCoK}
%k_{[n+1]}(\bm{x},\bm{x}') &=& k_{[n]}(\bm{x},\bm{x}') -c(\bm{x},\bm{x}',\bm{x}_{[n]}), \\
%c(\bm{x},\bm{x}',\bm{x}_{[n]})&=&\lambda_{[n+1,n+1]}(\bm{x})^T \left[k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]})+R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right] \lambda_{{[n+1,n+1]}}(\bm{x}'). \nonumber
%\end{eqnarray}
%\medskip
%
%We remark that, as noted in \cite{Chevalier2015} in the case of scalar-valued fields, these update formulae naturally 
%extend to Universal Kriging in second-order settings and apply without Gaussian assumption. 
%\end{comment}
%
%It turns out that the inclusion of several observations at a time (a batch) may be handled by the same equations as the 
%one for a single observation, provided some notation adjustments are made. This motivates the following.
Furthemore, boldface letters will be used to denote concatenated quantities corresponding to batches of observations.
Given a dataset consisting of  $q$ observations at spatial locations $\bm{\spatloc}=(\spatloc_1,\dots,\spatloc_q) \in \domain^q$ and response indices $\bm{\ell}=(\ell_1,\dots, \ell_q)\in \lbrace 1, ..., \no\rbrace^q$, %$i=1, ..., q$, 
we use the concatenated notation
\begin{align*}
\bm{x}:=%(\bm{\spatloc}, \bm{\ell}):= 
(x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
\end{align*}
We also compactly denote the field values at those different locations by
\begin{align*}
\gp[\bm{x}]:=
\left(\gp[\spatloc_1,\ell_1], ...,
\gp[\spatloc_q,\ell_{q}]\right) \in \mathbb{R}^{q}.
\end{align*}
%
For a second order random field $(Z_{\spatloc})_{\spatloc \in \domain}$ with mean $\mu$ and matrix covariance function $K$, $\mu$ is naturally extended to $\domain \in \lbrace 1, ..., \no\rbrace$ into a function of $x=(\spatloc, \ell)$ and is further straightforwardly vectorized into a function of $\bm{x}$. As for $K$, it induces a covariance kernel $k$ on the set of extended locations via  $k((\spatloc, \ell),(\spatloc', \ell'))=K(\spatloc, \spatloc')_{\ell, \ell'}$. In vectorized/batch form, $k(\bm{x}, \bm{x}')$ then amounts to a matrix with numbers of lines and columns the numbers of generalized locations in the batches $\bm{x}$ in $\bm{x}'$, respectively. Such vectorized quantities will turn out to be quite useful in order to arrive at embarrassingly simple expressions for co-Kriging equations below.  
%\textcolor{red}{(ADDENDUM, C.T.) In the same fashion, we will use $\mu(\bm{x})$ to denote the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ for the corresponding $q \times q$ covariance matrix.}

@@

Now given a GRF $\gp$ and observations of some of its components at some points in the domain, one can predict the value of the field at some unobserved location $\spatloc\in \domain$ by using the conditional mean of $\gp[\spatloc]$, conditional on the data. This coincides with the method called (co-)kriging, and (co-)kriging equations precisely tell us how to compute conditional means and covariances conditional on an arbitrary dataset.
We will here present a rather general form of cokriging, in the sense that it allows including several observations (batch) at a time and 
observations at a given location $u \in \domain$ may only include a subset of the components of $\gp[\spatloc]\in\mathbb{R}^{\no}$ 
(heterotopic).

Assuming that $n$ batches of observations are available, with respective sizes $q_1,\dots, q_n$, and that one wishes 
to predict $\gp[\bm{x}]$ for some batch of $q\geq 1$ generalized locations $\bm{x}$, %\equiv(\bm{\spatloc}, \bm{\ell})$, 
the (simple) cokriging mean then amounts to kriging with respect to a scalar-valued GRF indexed by 
$\domain\times \{1\dots,p\}$:  
%
\begin{equation}\label{eq:cokrig_mean}
\mu_{[n]}(\bm{x})=\mu(\bm{x})+\lambda_{[n]}(\bm{x})^T (\mathbf{z}_{[n]}-\mu(\bm{x})),
\end{equation}
where $\mu$ is $Z$'s initial mean function, $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n q_i$)-dimensional vector of observed (noisy) responses of $Z$ at all considered generalized locations, and $\lambda_{[n]}(\bm{x})$ is a vector of weights
equal to 
$$\left(k(\bm{x}_{[n]}, \bm{x}_{[n]})+\Delta_{[n]} \right)^{-1} k(\bm{x}_{[n]}, \bm{x})
$$
with $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$ and where $\Delta_{[n]}$ stands for the covariance matrix of Gaussian-distributed noise assumed to have affected measurements up to batch $n$. The matrix in parenthesis in the above equation will be assumed non-singular throughout the presentation. The co-kriging %(conditional) 
residual (cross-)covariance function (with respect to batches of generalized locations) can also be expressed in the same vein via
%
\begin{equation}\label{eq:cokrig_cov}
k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T 
\left(k(\bm{x}_{[n]}, \bm{x}_{[n]})+\Delta_{[n]} \right)
\lambda_{{[n]}}(\bm{x}').
\end{equation}

@@

Let us now consider the case where co-kriging prediction of $Z$ was made with respect to $n$ batches $\bm{x}_i$ of generalized locations, concatenated again within
$\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$, and one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
%It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
Thanks to our representation of co-kriging in terms of simple kriging with respect to generalized locations, a strightforward adaptation of the batch-sequential kriging update formulae from \citep{Chevalier.etal2013a} delivers that
% 
\begin{equation}\label{eq:meanCoK}
\mu_{[n+1]}(\bm{x})=\mu_{[n]}(\bm{x})+\lambda_{[n+1,n+1]}(\bm{x})^T (\mathbf{z}_{n+1}-\mu(\bm{x}_{n+1})),
\end{equation}
where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional sub-vector extracted from
$\lambda_{[n+1]}(\bm{x})$ that corresponds to the kriging weigths associated with the last $q_{n+1}$ responses when 
predicting at $\bm{x}$ relying on all measurements until batch $(n+1)$.
%\text{th}$ batch.
%, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
Similarly, the updated co-kriging residual (cross-)covariance function then writes
\begin{align}\label{eq:varCoK}
k_{[n+1]}(\bm{x},\bm{x}') & = k_{[n]}(\bm{x},\bm{x}')\\
 \nonumber - & \lambda_{[n+1,n+1]}(\bm{x})^T 
\left(k_{[n]}(\bm{x}_{n+1}, \bm{x}_{n+1})+\Delta_{n+1}\right)
%k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]})
\lambda_{{[n+1,n+1]}}(\bm{x}'),
\end{align}
%\medskip
%
Let us remark that, as noted in \citep{Chevalier2015} in the case of scalar-valued fields, these update formulae naturally 
extend to Universal Kriging in second-order settings and apply without Gaussian assumption. We will now see how the latter formulae are instrumental in deriving semi-analytical formulae for stepwise uncertainty reduction criteria for 
vector-valued random fields.

\subsection{UQ on ESs of multivariate GRFs}\label{sec:set_uq}
We now introduce quantities that allow UQ on the volume of the ES $\es$. Let $\mes$ be a 
(locally finite, Borel) measure  on $\domain$. We want to investigate the probability distribution 
of $\mes(\es)$ through its moments.
Centred moments may be computed using Proposition~\ref{propo1} developed in the appendix. 
In particular, the excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$ is an integral of the EP
\begin{equation*}
\begin{split}
\emv
&=\int_{\domain^2} \mathbb{P}\left(
\gp[u]\in T, \gp[v]\in T \right)
d\mes^{\otimes}(u, v)\\
&-\left( \int_{\domain} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
\end{split}
\end{equation*}
which in the excursion/sojourn case where $\T=(-\infty, t_1] \times
\dots \times (-\infty, t_{\no}]$ is
\begin{equation*}
\begin{split}
\emv
%\operatorname{Var}[\mes(\es)]
&=\int_{\domain^2}
\varPhi_{2\no}
\left(
(\bt, \bt); \mu((u,v)),
K((u,v),(u,v))
\right)
\
\mathrm{d}\mes^{\otimes} %\mes
%\productMeasure
(u,v)\\
&-\left( \int_{\domain} \varPhi_{\no}\left(\bt;\mu(u), K(u)\right) d\mes(u) \right)^2,
\end{split}
\end{equation*}
where $\varPhi_{\no}$ denotes the $\no$-variate Gaussian cumulative distribution function (CDF), which can be calculated using for instance the code of \cite{genz2009computation}.

%
Note that this quantity requires the solution of an integral over $\domain^2$. In 
contrast, and still like in the scalar-valued case, the IBV of \cite{bect2019} involves 
solely an integral on $\domain$ and can be expanded as 
\begin{equation*}
\begin{split}
\operatorname{IBV} %(\es) %;\mes)
&=\int_{\domain}
\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
d\mes(u) \\
&=\int_{\domain}
\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right)
-\left(\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right) \right)^2
\mathrm{d}\mes(u).
\end{split}
\end{equation*}
%

\subsection{Expected IBV and EMV}
\label{sec:eibv}
We compute the expected effect of the inclusion of new observations on the $\emv$ and $\ibv$ of the excursion set $\es$. Let us consider the same setting as in Eqn. \eqref{eq:meanCoK} and \eqref{eq:varCoK}, and let 
$\currentExp{.}$ and $\currentProba{.}$ denote conditional expectation
and probability conditional on the first $n$ batches of observations, respectively. We want to compute the effect of including a new set of observations at $\bm{x}_{n+1}$ on the $\emv$ and $\ibv$. We use $\IBV_{\stage}$ to denote $\IBV$ with respect to the conditional law $\mathbb{P}_{\stage}$.

@@

In order to study the effect of the inclusion of a new data point, we let
$
\currentIBV(\bm{x}; \bm{y})
$
denote the expected IBV under the current law of the field, conditioned on observing $\bm{y}$ at $\bm{x}$ (generalized, possibly batch observation). The expected effect of a new observation on the IBV is then captured by
\begin{equation}\label{def:eibv}
    \currentEIBV(\bm{x}):=\mathbb{E}_{\stage}\left[\mbox{IBV}(\bm{x}; \bm{Y})\right]
\end{equation}
where $\bm{Y}$ is distributed according to the current law of $Z_{\bm{x}}$ and with independent noise having covariance matrix $\Delta_n$.

@@

This integrand finally boils down to an evaluation of the $2\no$-dimensional Gaussian CDF by
Proposition~\ref{propo3}. Details are omitted here for brevity as we did not implement further
this criterion that requires numerical integration over $\domain^2$.
\end{propo}

We remark that 
Propositions \ref{propo_eibv} and \ref{propo_emv} are twofold generalizations of  results from \cite{chevalier2014fast}: they extend previous results to the multivariate setting and also allow for the inclusion of batch or heterotopic observations through the concept of generalized locations. 
A key element for the understanding of the propositions is that the conditional co-Kriging means entering in the EPs depend linearly on (batch) observations. The conditional equality expressions thus become linear combinations of Gaussian variables whose mean and covariance are easily calculated. 
Related closed-form solutions have been noted in similar contexts \citep{bhattacharjya2013value,stroh}, but not generalized to our situation with
random sets for vector-valued GRFs.

@@

We illustrate the EBV associated to different designs on a simple bivariate example. The example mimics our river plume application and hence the first and second component of the random field will be called \textit{temperature} and \textit{salinity} for illustrative purpose. Effects of the hyperparameters of the GRF prior on the excursion probabilities will also be studied. For simplicity, we begin with a \textit{pointwise} example, considering a single bivariate gaussian distribution (no spatiality).

@@

Say we want to study the excursion probability of a bivariate gaussian above some thresholds, where the thresholds are set equal to the mean; $\mu_1=t_1=5^o C$ for temperature and  $\mu_2=t_2=30$ g/kg for salinity, and we play with the temperature and salinity correlation and variances to study the effect on the EP and expected Bernoulli variance.

% To BE COMMENTED (TEMPORARILY?)
Fig. \ref{illus_bivarDens} shows contour plots of three different
densities with increasing correlation $\gamma$ between temperature and
salinity. 
\begin{figure}[h!] \centering
  \includegraphics[width=0.99\textwidth]{Figures/illus_bivar.eps}
  \caption{Density contour plots with different correlations between
    temperature and salinity. The densities have unit variance and the
    thresholds are identical to the mean values $5^o C$ and
    $30$ g/kg.}
\label{illus_bivarDens}
\end{figure}
The displayed densities have unit standard deviations for both
temperature and salinity, but we also study the effect of doubling the
standard deviations.
% END TO BE COMMENTED (TEMPORARILY?)

@@

We now study how the expected Bernoulli variance eq.\eqref{def:eibv} associated to data collection at a point changes if only one of the two components of the field is observed. We first draw a realization of the GRF defined above and use it as ground-truth to mimick the real data-collection process. A first bunch of observations are done at the locations depicted in grey, and the data is used to update the GRF model. We then consider the green triangle as a potential next observation locations and plot the expected Bernoulli variance reduction (at each grid point) that would result from observing only one component of the field (temperature or salinity), or both at that point, see Fig.\ref{fig:ebv_comp}.

@@

Note that plotting the Bernoulli variance reduction at each point might also be used to compare different data collection plans. For example, Fig. \ref{fig:ebv_north_vs_east} shows the expected Bernoulli variance reduction associated to a data collection plan along a straight upwards line (static north) and the one associated to a straight horizontal line (static east). Both expectations are computed according to the a priori distribution of the GRF (i.e. no observations have been included yet).

>>>>

%The methods we develop are fairly general, and not limited to the special case of two-dimensional random fields. Hence, the rest of this section will consider generic excursion sets of gaussian random fields with an arbitrary number of output dimensions.
% Say we have an underlying phenomenon that is modeled as a $\no$-variate gaussian random field $\gp$ on some domain $\domain$, 
For the general setting of a $\no$-variate random field, we are
interested in recovering the set of locations $\es$ in the domain for
which the components of $\gp$ lie in some set of specified values
$\T\subset \mathbb{R}^{\no}$; in other words \textit{the pre-image of
  $T$ by $\gp$}:
$$
\es:=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
$$
%
%Our goal here is to develop approaches to quantify and improve the
%characterization of uncertainties on $\es$. \textcolor{red}{Alread mention how observations are included.}

If we assume that $\gp$ has continuous trajectories 
%(almost surely \kc{This seems very handwavy; would delete}) 
and $T$ is closed, then $\es$ becomes a Random Closed Set
\citep{Molchanov2005} and concepts from the theory of random sets will
prove useful to study $\es$.
%under some measure defined on the domain.
%\textcolor{red}{Is this really needed?} 
Note that while some aspects of the developed approaches do not call
for a specific form of $\T$, we will often, for purposes of
simplicity, stay with the case of orthants
($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where
$t_1,\dots, t_{\no} \in \R$) as this will allow efficient calculation
of several key quantities. Note that changing some $\leq$ inequalities
to $\geq$ ones would lead to immediate adaptations.

@@

%Given the observations from $n$ batches, one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$. The mean update formulae is
%It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
% delivers that
% 
%\begin{equation}\label{eq:meanCoK}
%\mu_{[n+1]}(\bm{x})=\mu_{[n]}(\bm{x})+\lambda_{[n+1,n+1]}(\bm{x})^T (\mathbf{z}_{n+1}-\mu(\bm{x}_{n+1})),
%\end{equation}
%where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional sub-vector extracted from
%$\lambda_{[n+1]}(\bm{x})$ that corresponds to the Kriging weigths associated with the last $q_{n+1}$ responses when 
%predicting at $\bm{x}$ relying on all measurements until batch $(n+1)$.
%\text{th}$ batch.
%, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
%The updated co-Kriging residual cross-covariance function is
%\begin{eqnarray}\label{eq:varCoK}
%k_{[n+1]}(\bm{x},\bm{x}') &=& k_{[n]}(\bm{x},\bm{x}') -c(\bm{x},\bm{x}',\bm{x}_{[n]}), \\
%c(\bm{x},\bm{x}',\bm{x}_{[n]})&=&\lambda_{[n+1,n+1]}(\bm{x})^T \left[k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]})+R(\bm{x}_{[n]}, \bm{x}_{[n]}) \right] \lambda_{{[n+1,n+1]}}(\bm{x}'). \nonumber
%\end{eqnarray}
%\medskip
%
%We remark that, as noted in \cite{Chevalier2015} in the case of scalar-valued fields, these update formulae naturally 
%extend to Universal Kriging in second-order settings and apply without Gaussian assumption. 
%\end{comment}
%
%It turns out that the inclusion of several observations at a time (a batch) may be handled by the same equations as the 
%one for a single observation, provided some notation adjustments are made. This motivates the following.
Furthemore, boldface letters will be used to denote concatenated
quantities corresponding to batches of observations.  Given a dataset
consisting of
$q$ observations at spatial locations
$\bm{\spatloc}=(\spatloc_1,\dots,\spatloc_q) \in
\domain^q$ and response indices $\bm{\ell}=(\ell_1,\dots, \ell_q)\in
\lbrace 1, ..., \no\rbrace^q$, %$i=1, ..., q$,
we use the concatenated notation
\begin{align*}
\bm{x}:=%(\bm{\spatloc}, \bm{\ell}):= 
(x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
\end{align*}
We also compactly denote the field values at those different locations by
\begin{align*}
\gp[\bm{x}]:=
\left(\gp[\spatloc_1,\ell_1], ...,
\gp[\spatloc_q,\ell_{q}]\right) \in \mathbb{R}^{q}.
\end{align*}
%
For a second order random field $(Z_{\spatloc})_{\spatloc \in
  \domain}$ with mean $\mu$ and matrix covariance function $K$,
$\mu$ is naturally extended to $\domain \in \lbrace 1, ...,
\no\rbrace$ into a function of $x=(\spatloc,
\ell)$ and is further straightforwardly vectorized into a function of
$\bm{x}$. As for $K$, it induces a covariance kernel
$k$ on the set of extended locations via $k((\spatloc,
\ell),(\spatloc', \ell'))=K(\spatloc, \spatloc')_{\ell,
  \ell'}$. In vectorized/batch form, $k(\bm{x},
\bm{x}')$ then amounts to a matrix with numbers of lines and columns
the numbers of generalized locations in the batches
$\bm{x}$ in
$\bm{x}'$, respectively \kc{This last sentence doesn't parse.}. Such
vectorized quantities turn out to be useful in order to arrive at
simple expressions for co-Kriging equations below.
%\textcolor{red}{(ADDENDUM, C.T.) In the same fashion, we will use $\mu(\bm{x})$ to denote the $q$-dimensional vector corresponding to the mean at $\bm{x}$ and $k(\bm{x}, \bm{x})$ for the corresponding $q \times q$ covariance matrix.}

@@

Given a GRF $\gp$ and observations of some of its components at some
points in the domain, one can predict the value of the field at some
unobserved location $\spatloc\in \domain$ by using the conditional
mean of $\gp[\spatloc]$, conditional on the data \kc{...conditional
  mean on the data?}. This coincides with (co-)kriging equations,
which tell us precisely how to compute conditional means and
covariances conditional on an arbitrary dataset.  We will present a
general form of co-kriging, in the sense that it allows inclusion of
several observations (batch) at a time; observations at a given
location $u \in \domain$ may only include a subset of the components
of $\gp[\spatloc]\in\mathbb{R}^{\no}$ (heterotopic).

Assuming that $n$ batches of observations are available with sizes
$q_1,\dots, q_n$, and that one wishes to predict $\gp[\bm{x}]$ for
some batch of $q\geq 1$ generalized locations
$\bm{x}$, %\equiv(\bm{\spatloc}, \bm{\ell})$,
the simple cokriging mean then amounts to kriging with respect to a
scalar-valued GRF indexed by $\domain\times \{1\dots,p\}$:
%
\begin{equation}\label{eq:cokrig_mean}
\mu_{[n]}(\bm{x})=\mu(\bm{x})+\lambda_{[n]}(\bm{x})^T (\mathbf{z}_{[n]}-\mu(\bm{x})),
\end{equation}
where $\mu$ is $Z$'s initial mean function,
$\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n
q_i$)-dimensional vector of observed (noisy) responses of
$Z$ at all considered generalized locations, and
$\lambda_{[n]}(\bm{x})$ is a vector of weights equal to
$$\left(k(\bm{x}_{[n]}, \bm{x}_{[n]})+\Delta_{[n]} \right)^{-1} k(\bm{x}_{[n]}, \bm{x})
$$
with $\bm{x}_{[n]}=(\bm{x}_1,\dots,
\bm{x}_n)$ and where
$\Delta_{[n]}$ stands for the covariance matrix of
Gaussian-distributed noise assumed to have affected measurements up to
batch
$n$. The matrix in parenthesis will be assumed to be non-singular
throughout the presentation. The co-kriging %(conditional)
residual (cross-)covariance function (with respect to batches of
generalized locations) can also be expressed in the same vein via
%
\begin{equation}\label{eq:cokrig_cov}
k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T 
\left(k(\bm{x}_{[n]}, \bm{x}_{[n]})+\Delta_{[n]} \right)
\lambda_{{[n]}}(\bm{x}').
\end{equation}

@@

Let us now consider the case where a co-kriging prediction of $Z$ was
made with respect to $n$ batches $\bm{x}_i$ of generalized locations,
concatenated again within $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$,
and one wishes to update the prediction by incorporating a new vector
of observations $\mathbf{z}_{n+1}$ measured at a batch of
$q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
%It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
Thanks to our representation of co-kriging in terms of simple kriging
with respect to generalized locations, a strightforward adaptation of
the batch-sequential kriging update formulae from
\citep{Chevalier.etal2013a} suggests that
% 
\begin{equation}\label{eq:meanCoK}
\mu_{[n+1]}(\bm{x})=\mu_{[n]}(\bm{x})+\lambda_{[n+1,n+1]}(\bm{x})^T (\mathbf{z}_{n+1}-\mu(\bm{x}_{n+1})),
\end{equation}
where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional
sub-vector extracted from $\lambda_{[n+1]}(\bm{x})$ that corresponds
to the kriging weigths associated with the last $q_{n+1}$ responses
when predicting at $\bm{x}$ relying on all measurements until batch
$(n+1)$.
%\text{th}$ batch.
%, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
Similarly, the updated co-kriging residual (cross-)covariance function
then becomes
\begin{align}\label{eq:varCoK}
k_{[n+1]}(\bm{x},\bm{x}') & = k_{[n]}(\bm{x},\bm{x}')\\
 \nonumber - & \lambda_{[n+1,n+1]}(\bm{x})^T 
\left(k_{[n]}(\bm{x}_{n+1}, \bm{x}_{n+1})+\Delta_{n+1}\right)
%k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]})
\lambda_{{[n+1,n+1]}}(\bm{x}'),
\end{align}
%\medskip
%
As noted in \citep{Chevalier2015} in the case of scalar-valued fields,
these update formulae naturally extend to Universal Kriging in
second-order settings and apply without Gaussian assumptions. We will
now see how the latter formulae are instrumental in deriving
semi-analytical formulae for stepwise uncertainty reduction criteria
for vector-valued random fields.

\subsection{Uncertainty Quantification on ESs of multivariate GRFs}
\label{sec:set_uq}

We now introduce quantities that allow UQ on the volume of the ES
$\es$. Let $\mes$ be a (locally finite, Borel) measure on
$\domain$. We want to investigate the probability distribution of
$\mes(\es)$ through its moments.  Centred moments may be computed
using Proposition~\ref{propo1} developed in the appendix.  In
particular, the excursion measure variance
$\emv = \operatorname{Var}[\mes(\es)]$ is an integral of the excursion
probabilities (EP)

\begin{equation*}
\begin{split}
\emv
&=\int_{\domain^2} \mathbb{P}\left(
\gp[u]\in T, \gp[v]\in T \right)
d\mes^{\otimes}(u, v)\\
&-\left( \int_{\domain} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
\end{split}
\end{equation*}
which in the excursion/sojourn case where $\T=(-\infty, t_1] \times
\dots \times (-\infty, t_{\no}]$ is
\begin{equation*}
\begin{split}
\emv
%\operatorname{Var}[\mes(\es)]
&=\int_{\domain^2}
\varPhi_{2\no}
\left(
(\bt, \bt); \mu((u,v)),
K((u,v),(u,v))
\right)
\
\mathrm{d}\mes^{\otimes} %\mes
%\productMeasure
(u,v)\\
&-\left( \int_{\domain} \varPhi_{\no}\left(\bt;\mu(u), K(u)\right) d\mes(u) \right)^2,
\end{split}
\end{equation*}

where $\varPhi_{\no}$ denotes the $\no$-variate Gaussian cumulative
distribution function (CDF), which can be calculated, using for
instance, the code of \cite{genz2009computation}.

Note that this quantity requires the solution of an integral over
$\domain^2$. In contrast, and as in the scalar-valued case, the IBV of
\cite{bect2019} involves solely an integral on $\domain$ and can be
expanded as
\begin{equation*}
\begin{split}
\operatorname{IBV} %(\es) %;\mes)
&=\int_{\domain}
\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
d\mes(u) \\
&=\int_{\domain}
\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right)
-\left(\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right) \right)^2
\mathrm{d}\mes(u).
\end{split}
\end{equation*}
%

\subsection{Expected IBV and EMV}
\label{sec:eibv}

We compute the expected effect of the inclusion of new observations on
the $\emv$ and $\ibv$ of the excursion set $\es$. Let us consider the
same setting as in Eqn. \eqref{eq:meanCoK} and \eqref{eq:varCoK}, and
let $\currentExp{.}$ and $\currentProba{.}$ denote conditional
expectation and probability conditional on the first $n$ batches of
observations, respectively. We want to compute the effect of including
a new set of observations at $\bm{x}_{n+1}$ on the $\emv$ and
$\ibv$. We use $\IBV_{\stage}$ to denote $\IBV$ with respect to the
conditional law $\mathbb{P}_{\stage}$.

@@

In order to study the effect of the inclusion of a new data point, we
let $ \currentIBV(\bm{x}; \bm{y}) $ denote the expected IBV under the
current law of the field \kc{Is this a legitimate expression 'law of
  the field'?}, conditioned on observing $\bm{y}$ at $\bm{x}$
(generalized, possibly batch observation). The expected effect of a
new observation on the IBV is then shown by
\begin{equation}\label{def:eibv}
    \currentEIBV(\bm{x}):=\mathbb{E}_{\stage}\left[\mbox{IBV}(\bm{x}; \bm{Y})\right]
\end{equation}
where $\bm{Y}$ is distributed according to the current law of
$Z_{\bm{x}}$ and with independent noise having covariance matrix
$\Delta_n$.

@@

This integrand finally boils down to an evaluation of the
$2\no$-dimensional Gaussian CDF by Proposition~\ref{propo3}. Details
are omitted here for brevity as we do not implement this criterion
further and which requires numerical integration over $\domain^2$.
\end{propo}

We remark that Propositions \ref{propo_eibv} and \ref{propo_emv} are
twofold generalizations of results from \cite{chevalier2014fast}: they
extend previous results to the multivariate setting and also allow for
the inclusion of batch or heterotopic observations through the concept
of generalized locations.  A key element for understanding of these
propositions is that the conditional co-Kriging means entering in the
EPs depend linearly on (batch) observations. The conditional equality
expressions thus become linear combinations of Gaussian variables
whose mean and covariance are easily calculated.  Related closed-form
solutions have been noted in similar contexts
\citep{bhattacharjya2013value,stroh}, but not generalized to our
situation with random sets for vector-valued GRFs.

@@

We illustrate the EBV associated with different designs on a simple
bivariate example. This mimics our river plume application and hence
the first and second component of the random field will be called
\textit{temperature} and \textit{salinity} for illustrative
purpose. Effects of the hyperparameters of the GRF prior on the
excursion probabilities will also be studied. For simplicity, we begin
with a \textit{pointwise} example, considering a single bivariate
gaussian distribution (i.e. no spatiality).

@@

Say we want to study the excursion probability of a bivariate gaussian
above some threshold, where the thresholds are set equal to the mean;
$\mu_1=t_1=5^o C$ for temperature and $\mu_2=t_2=30$ g/kg for
salinity, and we play with the temperature and salinity correlation
and variances to study the effect on the EP and expected Bernoulli
variance.

\begin{figure}[!b] 
\centering
\includegraphics[width=0.99\textwidth]{Figures/illus_bivar.eps}
\caption{Density contour plots with different correlations between
  temperature and salinity. The densities have unit variance and the
  thresholds are identical to the mean values $5^o C$ and
  $30$ g/kg.}
\label{illus_bivarDens}
\end{figure}

Fig. \ref{illus_bivarDens} shows contour plots of three different
densities with increasing correlation $\gamma$ between temperature and
salinity. The displayed densities have unit standard deviations for
both temperature and salinity, but we also study the effect of
doubling the standard deviations.

@@

We now study how the expected Bernoulli variance eq.\eqref{def:eibv}
associated with data collection at a point changes if only one of the
two components of the field is observed. We first draw a realization
of the GRF defined above and use it as ground-truth to mimick the real
data-collection process. A first set of observations are done at the
locations depicted in grey (see Fig.~\ref{fig:ebv_comp}), and the data
is used to update the GRF model. We then consider the green triangle
as a potential next observation location and plot the expected
Bernoulli variance reduction (at each grid point \kc{I don't recall we
  have been explicit about grids before this mention. It might be
  useful to recall that here.}) that would result from observing only
one component of the field (temperature or salinity), or both at that
point.

@@

Note that plotting the Bernoulli variance reduction at each point
might also be used to compare different data collection plans. For
example, Fig. \ref{fig:ebv_north_vs_east} shows the expected Bernoulli
variance reduction associated with a data collection plan along a
vertical line (static north) and one associated with a horizontal
(static east). Both expectations are computed according to the
a-priori distribution of the GRF (i.e. no observations have been
included yet).
